---
title: "2.1 Varroa Infestation"
author: "Dan Aurell"
date: "2023-05-22"
output:   
  html_document: 
    keep_md: yes
---

# Setup

```{r}
library(MASS)
library(tidyverse)
library(lme4)
library(boot)
library(emmeans)
library(broom)
library(DHARMa)
```

Vector of queen events
```{r}
# From "collated qlessness" sheet

# Any colonies that became queenless between day 0 and 42
qless <- c("D8", "RC1", "D21", "D24", "C6", "BC15")

```

```{r}
datum <- read.csv("../2 Analysis/data_prepped/datum_prepped.csv")
```

Check treatment group of the colonies that were removed
```{r}
datum %>% 
  filter(days == 0, col_no %in% qless)

# 4 were in Apivar
# 2 were in Combination
```


```{r, include=FALSE}
datum$trt <- factor(datum$trt , levels=c("Apivar", "Apiguard_Api", "AEC"))
# datum$trt <- factor(datum$trt , levels=c("Apiguard_Api", "AEC", "Apivar"))

# So that days is treated as categorical in models
datum$days <- factor(datum$days)
```

Data prep for phor, infest, bees modeling
```{r}
tempdatum <- datum %>% 
  filter(days %in% c(0,21,42),
         !(col_no %in% qless)
         )

# Simple sample size calculation to respond to reviewer
# These were manually entered into table generated in the Plotting folder

# Relevant variables with same n: trt days mites_recov_3_washes bees

tempdatum %>% 
  group_by(days, trt) %>% 
  summarize(n= n())


# Different n: infest_200
tempdatum %>% 
  filter(!is.na(infest_200)) %>% 
  group_by(days, trt) %>% 
  summarize(n= n())



```

Quick descriptive stats
```{r}
tempdatum %>%
 summarise(
 mean = mean(est_bee_num_by_wt),
 mini = min(est_bee_num_by_wt),
 maxi = max(est_bee_num_by_wt),
 variance = var(est_bee_num_by_wt),
 sd = sqrt(variance)
 )


```




# Treatment Effects on Varroa Infestation
## Varroa on adult bees

### Poisson vs. Negative Binomial

Checking for overdispersion - 
```{r}
dispersionstats <- datum %>%
    filter(days %in% c(0,21,42),
         !(col_no %in% qless)) %>% 
 summarise(
 means = mean(mites_recov_3_washes),
 variances = var(mites_recov_3_washes),
 ratio = variances/means)

# Variance-to-mean ratio of 22.5; use negative binomial distribution instead of Poisson
```


Fit Poisson and negative binomial models
```{r}
m.phor.pois1 <- glmer(mites_recov_3_washes ~ 
                     trt + 
                     days + 
                     trt:days + 
                     (1 | col_no), 
                   data = tempdatum, family = poisson)


m.phor.nb1 <- glmer.nb(mites_recov_3_washes ~ 
                     trt + 
                     days + 
                     trt:days + 
                     (1 | col_no), 
                   data = tempdatum)

summary(m.phor.pois1)
summary(m.phor.nb1)

```


Comparing nb to poisson with deviance F-test
```{r}
# Test whether nb is a stat sig better fit than poisson:
anova(m.phor.nb1, m.phor.pois1)
# Yes, highly significantly better (p<2.2e-16)

summary(m.phor.pois1)$coefficients
summary(m.phor.nb1)$coefficients
# The p values for my predictors went up, but the fit is way better
  # This makes sense... Poisson is not conservative enough with overdispersed data
# The parameter estimates stay pretty much the same. Good.
```

### Fit model that uses number of bees in sample as the offset

```{r}
# A priori model
m.phor.nb1 <- glmer.nb(mites_recov_3_washes ~ 
                     trt + 
                     days + 
                     trt:days + 
                     (1 | col_no), 
                   data = tempdatum)
summary(m.phor.nb1)

# Take into account the number of bees per sample

m.phor <- glmer.nb(mites_recov_3_washes ~ 
                     trt + 
                     days + 
                     trt:days + 
                     offset(log(est_bee_num_by_wt_floor)) +
                     (1 | col_no),
                   data = tempdatum)
summary(m.phor) 
# I need to provide a value of the offset when making predictions from the model


# Test significance of trt:days interaction

## Fit model that omits trt:days interaction

m.phor.0 <- glmer.nb(mites_recov_3_washes ~ 
                     trt + 
                     days + 
                     # trt:days + 
                     offset(log(est_bee_num_by_wt_floor)) +
                     (1 | col_no),
                   data = tempdatum)
summary(m.phor.0)

anova(m.phor, m.phor.0) 
# There was a significant interaction between day and treatment (p=2.251e-05) when tested by ANOVA. Therefore we analyzed each date separately.

# Final model is m.phor ".f" means "final"
m.phor.f <- m.phor
```

### Joint tests of phor within day (Like ANOVA of phor within day)
Added step based on reviewer comment

```{r}
#joint_tests(m.phor, by = "days") 
# Duplicate row names not allowed error

#joint_tests(m.phor, offset = log(100))
# Error in .find.by.rows(args, by) : 'by' variables are not all in the grid

#joint_tests(m.phor, offset = log(100), by = "days")
# Error in .find.by.rows(args, by) : 'by' variables are not all in the grid

# Used the joint_tests documentation to recreate the individual steps of what joint_tests does, adding in the offset at the appropriate step: https://cran.r-hub.io/web/packages/emmeans/emmeans.pdf

emm.phor.f <- emmeans(m.phor.f, pairwise ~ trt | days)
con.phor.f <- contrast(emm.phor.f, offset = log(100))
test(con.phor.f, joint = TRUE, by = "days")
# On day 0, no significant differences between treatments (F ratio 2,Inf = 0.187, P=0.8291)
# On day 21, significant differences between treatments (F ratio 2,Inf = 9.388, P=0.0001)
# On day 42, no significant differences between treatments (F ratio 2,Inf = 2.542, P=0.0787)


```





## Infestation of Brood

Checking for overdispersion
```{r}
dispersionstats2 <- datum %>%
    filter(days %in% c(0,21,42),
         !(col_no %in% qless)) %>% 
 summarise(
 means = mean(infest_200, na.rm = T),
 variances = var(infest_200, na.rm = T),
 ratio = variances/means)

# Variance-to-mean ratio of 14.2; use negative binomial distribution instead of Poisson
```

```{r}
# a priori model
m.infest.nb1 <- glmer.nb(infest_200 ~ 
                     trt + 
                     days + 
                     trt:days + 
                     (1 | col_no),
                   data = tempdatum)
summary(m.infest.nb1)

m.infest.nb0 <- glmer.nb(infest_200 ~ 
                     trt + 
                     days + 
                     # trt:days + 
                     (1 | col_no),
                   data = tempdatum)

anova(m.infest.nb1, m.infest.nb0)
# We found a significant interaction of treatment × experiment day on infestation rate of capped
# worker brood (ChiSq 4=9.74, P=0.045)


# Final model
m.infest.f <- m.infest.nb1

```


### Joint tests of infest within day (Like ANOVA of infest within day)
Added step based on reviewer comment

```{r}
joint_tests(m.infest.f) # Does not give a duplicate row names error

joint_tests(m.infest.f, by = "days") 

# Testing that this method (used for phor) gives same result
emm.infest.f <- emmeans(m.infest.f, pairwise ~ trt | days)
con.infest.f <- contrast(emm.infest.f, offset = log(100))
test(con.infest.f, joint = TRUE, by = "days")
# Yes, exactly the same output. This confirms the workflow used above for phor

# On day 0, no significant differences (F ratio 2,Inf=0.004 , P=0.9957)
# On day 21, yes significant differences (F ratio 2,Inf= 5.678, P=0.0034)
# On day 42, no significant differences (F ratio 2,Inf= 0.427, P=0.6525)
```



# Summarize models for text, tables, and plotting

```{r}
# Estimated marginal means (and CIs) for text, tables, and plotting
# For the models which included an offset, report the emm's and CIs for a hypothetical sample of 100 bees
# This method is based on the emmeans documentation at: https://cran.r-project.org/web/packages/emmeans/vignettes/sophisticated.html#offsets

emplot_phor <- emmip(m.phor.f, trt~days, type = "response", CIs = TRUE, plotit = FALSE, offset = log(100))
emplot_phor$days <- as.numeric(as.character(emplot_phor$days))
emplot_phor$trt <- factor(emplot_phor$trt, levels = c("Apivar", "AEC", "Apiguard_Api"))
levels(emplot_phor$trt) # check order of factors

emplot_infest <- emmip(m.infest.f, trt~days, type = "response", CIs = TRUE, plotit = FALSE)
emplot_infest$days <- as.numeric(as.character(emplot_infest$days))
emplot_infest$trt <- factor(emplot_infest$trt, levels = c("Apivar", "AEC", "Apiguard_Api"))
levels(emplot_infest$trt) # check order of factors
# This is simply the number of infested cells per 200 worker cells
# Now scale this down to infested per 100 worker cells
emplot_infest$yvar <- emplot_infest$yvar/2
emplot_infest$LCL <- emplot_infest$LCL/2
emplot_infest$UCL <- emplot_infest$UCL/2


```


# Save these summarized tables of model estimates to their own CSVs

```{r}
write.csv(emplot_phor, "../3 Plotting/data_analyzed/emplot_phor.csv")
write.csv(emplot_infest, "../3 Plotting/data_analyzed/emplot_infest.csv")
```

# Post-hoc comparisons

```{r}
# Phor within day
emmeans(m.phor.f, pairwise ~ trt | days, type = "response")$contrasts
confint(emmeans(m.phor.f, pairwise ~ trt | days, type = "response"))$contrasts

# Confirm that emmeans is using asymptotic degrees of freedom
emmeans(m.phor.f, pairwise ~ trt | days, type = "response")$contrasts
emmeans(m.phor.f, pairwise ~ trt | days, type = "response", lmer.df = "asymptotic")$contrasts
# Yes, same results

# Does Satterthwaite method work? No, defaults to asymptotic
emmeans(m.phor.f, pairwise ~ trt | days, type = "response")$contrasts
emmeans(m.phor.f, pairwise ~ trt | days, type = "response", lmer.df = "satterthwaite")$contrasts

# Infest within day
emmeans(m.infest.f, pairwise ~ trt | days, type = "response")$contrasts
confint(emmeans(m.infest.f, pairwise ~ trt | days, type = "response"))$contrasts

```


