---
title: "3_colony_strength_2023-05-22"
author: "Dan Aurell"
date: "2023-05-22"
output:   
  html_document: 
    keep_md: yes
---

# Last modified

# Setup

```{r}
library(MASS)
library(tidyverse)
library(lme4)
library(boot)
library(emmeans)
library(broom)
library(DHARMa)
```

Vector of queen events
```{r}
# From "collated qlessness" sheet

# Any colonies that became queenless between day 0 and 42
qless <- c("D8", "RC1", "D21", "D24", "C6", "BC15")

```

```{r}
datum <- read.csv("../2 Analysis/data_prepped/datum_prepped.csv")
```

```{r, include=FALSE}
datum$trt <- factor(datum$trt , levels=c("Apivar", "Apiguard_Api", "AEC"))
# datum$trt <- factor(datum$trt , levels=c("Apiguard_Api", "AEC", "Apivar"))

# So that days is treated as categorical in models
datum$days <- factor(datum$days)
```

Data prep for phor, infest, bees modeling
```{r}
tempdatum <- datum %>% 
  filter(days %in% c(0,21,42),
         !(col_no %in% qless)
         ) %>%
  mutate(tot = wk_cells_infested_estimate + phoretic_mites_num_estimate,
           perc_of_mitepop_wkbrd = 100*wk_cells_infested_estimate/tot
           )
```


# Treatment Effects on Colony Strength
## Adult Bee Population

### Backwards variable testing starting from a priori model

```{r}
# A priori model (REML)
m.bee.a <- lmer(bees_frames ~ 
                 trt + 
                 days + 
                 trt:days +
                 (1 | col_no), data = tempdatum)
summary(m.bee.a) # 119 obs, 41 groups

# I think the inclusion of the random effect of colony number is essential to avoid pseudoreplication, but Zach seemed to think it may be a case of overfitting

# Let's test out an alternate model
m.bee.alm <- lm(bees_frames ~ 
                 trt + 
                 days + 
                 trt:days, 
                data = tempdatum)
summary(m.bee.alm) # 110 degrees of freedom

# What's our true number of samples for the effect of time, the effect of treatment, and the time x treatment interaction? Refer back to Steury notes

# Temporal pseudoreplication
# Repeated measures design
# lm(Height ~ Age + Sex) # Pseudoreplicated with respect to Sex, but not Age
# Interactions not discussed
# Relevant to winter trial - Usually you don’t have to include random effects but doing so lets you incraese your power



# Check deletion of interaction (ML)
m.bee.a.ML <- lmer(bees_frames ~ 
                 trt + 
                 days + 
                 trt:days +
                 (1 | col_no), data = tempdatum, REML = F)

m.bee.a0.ML <- lmer(bees_frames ~ 
                 trt + 
                 days + 
                 # trt:days +
                 (1 | col_no), data = tempdatum, REML = F)

anova(m.bee.a.ML, m.bee.a0.ML)
# The interaction of treatment and experiment day on adult bee population was significant (ChiSq[4]=12.097; P=0.01664) 
```

### Forward variable selection:
```{r}
# Check addition of a Varroa effect
m.bee.1.ML <- lmer(bees_frames ~ 
                 trt + 
                 days + 
                 trt:days +
                 pre_phor2 + 
                 (1 | col_no), data = tempdatum, REML = F)

anova(m.bee.1.ML, m.bee.a.ML)
# the model fit was significantly improved by including the effect of day-0 infestation rate on bee population (ChiSq=25.04; P=5.615e-07)

# Check addition of a Varroa:days effect
m.bee.2.ML <- lmer(bees_frames ~ 
                 trt + 
                 days + 
                 trt:days +
                 pre_phor2 + 
                 pre_phor2:days +
                 (1 | col_no), data = tempdatum, REML = F)

anova(m.bee.1.ML, m.bee.2.ML)
# The interaction of day-0 infestation rate and day on bee population also significantly improved model fit (ChiSq=7.8802, P=0.01945)

# Final model
m.bee.2 <- lmer(bees_frames ~ 
                 trt + 
                 days + 
                 trt:days +
                 pre_phor2 + 
                 pre_phor2:days +
                 (1 | col_no), data = tempdatum)

summary(m.bee.2)

m.bee.f <- m.bee.2

```


Interpret model
```{r}
summary(m.bee.f)
confint(m.bee.f)

```

### ANOVA of bees_frames within day

```{r}
joint_tests(m.bee.f, by = "days")

# On day 0, no significant differences (F ratio 2,71.27=0.325, P=0.7235)
# On day 21, yes significant differences (F ratio 2,71.27=5.192, P=0.0079)
# On day 42, yes significant differences (F ratio 2,77.46=4.516, P=0.0140)
```


On day X, yes/no significant differences between treatments (F = , P= )

```{r}
m.bee_day0 <- lm(bees_frames ~ trt + pre_phor2,
                 data = tempdatum, subset = days == 0)
summary(aov(m.bee_day0))

# On day 0, no significant differences between treatments (F 2,37 = 0.310, P=0.7353)

m.bee_day21 <- lm(bees_frames ~ trt + pre_phor2,
                 data = tempdatum, subset = days == 21)
summary(aov(m.bee_day21))

# On day 21, yes significant differences between treatments (F 2,37= 4.124, P= 0.0242)

m.bee_day42 <- lm(bees_frames ~ trt + pre_phor2,
                 data = tempdatum, subset = days == 42)
summary(aov(m.bee_day42))


# On day 42, no significant differences between treatments (F 2,33= 1.891, P= 0.167)
  # This is a bit surprising since it's at odds with the emmeans analysis
```



## Population of capped worker brood

```{r}
tempdatum <- datum %>% 
  filter(days %in% c(0,10,21,42),
         !(col_no %in% qless)
         )

tempdatum %>% 
  group_by(days, trt) %>% 
  summarize(n= n())
```

### Backward variable testing

```{r}
# A priori model (REML)
m.brood.a <- lmer(wk_frames ~ 
                   trt + 
                   days + 
                   trt:days + 
                  (1 | col_no), 
                 data = tempdatum)
summary(m.brood.a)


# Check deletion of interaction (ML)
m.brood.a0 <- lmer(wk_frames ~ 
                   trt + 
                   days + 
                   # trt:days + 
                  (1 | col_no), 
                 data = tempdatum)

anova(m.brood.a, m.brood.a0) # No need any more to specify ML models

# We did not find a significant interaction between treatment and days on worker brood population (ChiSq[6]=10.766; P=0.09588), but because of our experiment designed to examine this effect, we retained the interaction in the model anyway.

# Check deletion of trt, then days

m.brood.a00 <- lmer(wk_frames ~ 
                   # trt + 
                   days + 
                   # trt:days + 
                  (1 | col_no), 
                 data = tempdatum)

anova(m.brood.a0, m.brood.a00)
# Effect of trt not significant (Chisq[2] = 1.84, P=0.3994)



# Check deletion of days
m.brood.a000 <- lmer(wk_frames ~ 
                   trt + 
                   # days + 
                   # trt:days + 
                  (1 | col_no), 
                 data = tempdatum)
# DAN Singular

anova(m.brood.a0, m.brood.a000)
# Significant effect of days (Chisq[3]=192.35, P<2.2e-16).
```

### Forward variable selection

```{r}
# Check addition of a Varroa effect
m.brood.1 <- lmer(wk_frames ~ 
                   trt + 
                   days + 
                   trt:days + 
                   pre_phor2 +
                  (1 | col_no), 
                 data = tempdatum)

anova(m.brood.a, m.brood.1)
# The addition of Varroa was highly significant (ChiSq[1]=13.767, P=0.000207)


# Check addition of a Varroa:days effect
m.brood.2 <- lmer(wk_frames ~ 
                   trt + 
                   days + 
                   trt:days + 
                   pre_phor2 +
                   pre_phor2:days +
                  (1 | col_no), 
                 data = tempdatum)
anova(m.brood.2, m.brood.1)
# The addition of Varroa:days was not significant (ChiSq[3]=6.926, P=0.07429)

# Now that have added a Varroa effect, test deletion of the treatment:days interaction backwards

m.brood.3 <- lmer(wk_frames ~ 
                   trt + 
                   days + 
                   # trt:days + 
                   pre_phor2 +
                  (1 | col_no), 
                 data = tempdatum)

anova(m.brood.3, m.brood.1)
# Also in the presence of the pre_Varroa effect, there is a P=0.096 effect of


# Check that the effect estimate of trt:time is robust to the addition of the pre_Varroa effect
s.1 <- summary(m.brood.a)
s.2<- data.frame(s.1$coefficients)
s.3 <- data.frame(confint(m.brood.a))[3:14,]
s.4 <- cbind(s.2, s.3)

# Now estimates for model including the pre_Varroa effect
s.5 <- summary(m.brood.1)
s.6 <- data.frame(s.5$coefficients)[-7,]

# Bind together
s.7 <- cbind(s.4, s.6)

# The estimates of treatment effect are remarkably similar with the addition of pre_Varroa as a predictor of brood area. Only the new estimate of the intercept moved outside the bounds of the original 95% CI, and all the others stayed remarkably similar.

anova(m.brood.1)

```

```{r}
# Final model

# Preferred model: m.brood.1.ML
# Make REML version
m.brood.1 <- lmer(wk_frames ~ 
                   trt + 
                   days + 
                   trt:days + 
                   pre_phor2 +
                   # pre_phor2:days +
                  (1 | col_no), 
                 data = tempdatum)

summary(m.brood.1)
confint(m.brood.1)

# Final model is m.brood.1
m.brood.f <- m.brood.1
```



# Look at residuals of preferred models

```{r}
hist(resid(m.bee.f)) # Not too off from normal dist
hist(resid(m.brood.f)) # Looks pretty normal
```

### No need to do ANOVA of wk_frames within day
- Because the interaction wasn't significant


# Tidy the output of models
To be able to report predicted response

```{r}
# Bees

emres_bees<- emmeans(m.bee.f, pairwise ~ trt | days, type = "response")$emmeans %>% 
  tidy()

emmeans(m.bee.f, pairwise ~ trt | days, type = "response")$contrasts
confint(emmeans(m.bee.f, pairwise ~ trt | days, type = "response"))$contrasts


## Day 42 frames of bees estimates
### AEC 4.292991
### Apivar 2.912365
### Combination 2.951125

## What percentage of bees were lacking in Apivar?
100*(4.292991-2.912365)/4.292991 #32%

# 1.38 missing frames of bees. Convert to number of bees
# Beebook reference
2430*1.38 # 3353

## What percentage of bees were lacking in Combination?
100*(4.292991-2.951125)/4.292991 # 31%

## Combination missing 1.34 frames of bees. Convert to number of bees
## Beebook conversion
2430*1.34 # 3256 bees missing


## Combination missing how many frames of brood on Day 10?
1.2-0.73 # 0.47 frames of brood


880*3.8*2 # 6688 cells of worker brood per frame

## Combination missing 0.47 frames of brood on Day 10. Convert to number of bees
6688*0.47 #3143 brood missing








## Over time
emmeans(m.bee.f, pairwise ~ days | trt, type = "response")$contrasts

# Bees a priori

emres_bees_apriori<- emmeans(m.bee.a, pairwise ~ trt | days, type = "response")$emmeans %>% 
  tidy()

emmeans(m.bee.a, pairwise ~ trt | days, type = "response")$contrasts
confint(emmeans(m.bee.a, pairwise ~ trt | days, type = "response"))$contrasts

# Brood



emres_brood<- emmeans(m.brood.f, pairwise ~ trt | days, type = "response")$emmeans %>% 
  tidy()

emmeans(m.brood.f, pairwise ~ trt | days, type = "response")$contrasts
confint(emmeans(m.brood.f, pairwise ~ trt | days, type = "response"))$contrasts
# Note that we have a significant comparison here even though the inclusion of day:trt effect in the overall model was not significant.
```

# Summarize models for plotting

```{r}
emplot_bees <- emmip(m.bee.f, trt~days, type = "response", CIs = TRUE, plotit = FALSE)
emplot_bees$days <- as.numeric(as.character(emplot_bees$days))
emplot_bees$trt <- factor(emplot_bees$trt, levels = c("Apivar", "AEC", "Apiguard_Api"))
levels(emplot_bees$trt) # check order of factors

emplot_bees_apriori <- emmip(m.bee.a, trt~days, type = "response", CIs = TRUE, plotit = FALSE)
emplot_bees_apriori$days <- as.numeric(as.character(emplot_bees_apriori$days))
emplot_bees_apriori$trt <- factor(emplot_bees_apriori$trt, levels = c("Apivar", "AEC", "Apiguard_Api"))
levels(emplot_bees_apriori$trt) # check order of factors

emplot_brood <- emmip(m.brood.f, trt~days, type = "response", CIs = TRUE, plotit = FALSE)
emplot_brood$days <- as.numeric(as.character(emplot_brood$days))
emplot_brood$trt <- factor(emplot_brood$trt, levels = c("Apivar", "AEC", "Apiguard_Api"))
levels(emplot_brood$trt) # check order of factors
```


# Save these summarized tables of model estimates to their own CSVs

```{r}
write.csv(emplot_bees, "../3 Plotting/data_analyzed/emplot_bees.csv")
write.csv(emplot_brood, "../3 Plotting/data_analyzed/emplot_brood.csv")
```

