---
title: "1.1 Data Preparation"
author: "Dan Aurell"
date: "2023-04-05"
output:   
  html_document: 
    keep_md: yes
---

```{r}
library(MASS)
library(tidyverse)
```

# Load data, make simpler dataset --------

Rename
```{r}
datum_orig <- read.csv("./data/data9 flat v2.csv")

datum_orig <- datum_orig %>% rename(
  bees = bees_num,
  pattern = brd_pattern,
  phor = perc_phoretic_mites_by_weight,
  infest_200 = infested_in_first_200_cells_num,
  trt = treatment,
  wk_cells = wrk_capbrd_cells_num,
  fob_at_all = fob_covered_at_all,
  fob_top_bip = top_of_box_bip_style_fob
  
)
```


Here I change the units of a few columns and select the columns of interest
```{r}
datum <- datum_orig %>% 
  mutate(
    days = ifelse(time_point == 1, 0, 
                     ifelse(time_point == 10, 10, 
                            ifelse(time_point == 17, 21, 
                                   ifelse(time_point == 22, 42, 
                                          ifelse(time_point == 34, 138, NA))))),
    est_bee_num_by_wt_floor = floor(est_bee_num_by_wt),
    bees_without_mites = est_bee_num_by_wt_floor-mites_recov_3_washes,
    bees_frames = bees_tot_sides/2, # now expressed in frames
    wk_frames = wk_capbrd_tot_sides/2, # now expressed in frames
    honey_frames = revised_honey_tot_sides/2, # now expressed in frames
    pollen_frames = revised_pollen_tot_sides/2, # now expressed in frames
    phor2 = 100*mites_recov_3_washes/est_bee_num_by_wt, # make unrounded variable for phor
    cap_brd_indiv = dr_capbrd_cells_num + wk_cells, 
    indiv_total = cap_brd_indiv + bees, # Make a composite variable of both brood and bees
    indiv_capbrd_pct = 100*cap_brd_indiv/indiv_total,
    indiv_adult_pct = 100*bees/indiv_total,
    phoretic_mites_num_estimate = floor(phoretic_mites_num_estimate),
    wk_cells_infested_estimate = floor(wk_cells_infested_estimate),
    dr_cells_infested_estimate = floor(dr_cells_infested_estimate),
    mites_total_estimate = floor(mites_total_estimate),
    ) %>% 
  select(
    col_no,
    trt,
    time_point,
    date,
    days,
    q_status,
    q_event,
    pattern,
    notes,
    est_bee_num_by_wt,
    est_bee_num_by_wt_floor,
    mites_recov_3_washes,
    bees_without_mites,
    phor2, # Use unrounded column
    infest_200,
    dwv_num_sides_obs,
    dwv_tf,
    cdb_num_sides_obs,
    bees,
    wk_cells,
    dr_capbrd_cells_num,
    phoretic_mites_num_estimate,
    wk_cells_infested_estimate,
    dr_cells_infested_estimate,
    mites_total_estimate,
    cap_brd_indiv,
    indiv_total,
    indiv_adult_pct,
    indiv_capbrd_pct,
    bees_frames,
    wk_frames,
    honey_frames,
    pollen_frames,
    fob_at_all,
    fob_top_bip,
    col_hist
  ) %>%
  # Filter to include only the time points where we have data
  filter(days %in% c(0, 10, 21, 42, 138),
         (col_no != "C15" | days != "42") # Remove row for a deadout on Day 42
         )
datum[sapply(datum, is.infinite)] <- NA # To remove Inf as value in cells
datum[sapply(datum, is.nan)] <- NA


```


Vector of queen events
```{r}
# From "collated qlessness" sheet

# Any colonies that became queenless between day 0 and 42
qless <- c("D8", "RC1", "D21", "D24", "C6", "BC15")

```


How many colonies in each treatment group came from the repeated-OA vapor trial?
```{r}
datum %>% 
  filter(col_hist == "Christian OA")

datum %>% 
  filter(!(col_no %in% qless),
         col_hist == "Christian OA" 
  )
```

Curiosity: bee-brood ratio
```{r}
datum %>% 
  filter(!col_no %in% qless, days == 0) %>%
  summarise(
    mean_bees = mean(bees),
    mean_wk_capped = mean(wk_cells)
  )

# bees: 10309
# capped worker: 17345
# all worker: 17345*21/12 = 30353
# suitable: 30353/21 = 1445
# Assume 100 mg per bee, so 10,000 bees/ kg
# kg bees = 10309/10000 = 1.03
# cells suitable per kg bees = 1445/1.03 = 1402
  # Way extrapolates data from Boot. This would be a super short phoretic period with say 0.5 or more probability of invasion each day


```


```{r, include=FALSE}
datum$trt <- factor(datum$trt , levels=c("Apivar", "Apiguard_Api", "AEC"))
# datum$trt <- factor(datum$trt , levels=c("Apiguard_Api", "AEC", "Apivar"))

# So that days is treated as categorical in models
datum$days <- factor(datum$days)
```



# CALCULATE DAY-0 VARIABLES --------------

### Make new pre-treatment columns

Make a column of pre-treatment values for each variable of interest.
In order to do ANCOVA-POST if I wish to - and to include as predictors in mixed modeling:

```{r}
# Extract row based on colony number of row i and day = 0
# Write the value of the column of interest (say bees to start with) into a pre_bees column

#    pattern
for(i in 1:nrow(datum)) {
  datum$pre_pattern[i] <- datum[which(datum$days == 0 & datum$col_no == datum$col_no[i]), ]$pattern
}

#    phor2
for(i in 1:nrow(datum)) {
  datum$pre_phor2[i] <- datum[which(datum$days == 0 & datum$col_no == datum$col_no[i]), ]$phor2
}

#    infest_200
for(i in 1:nrow(datum)) {
  datum$pre_infest_200[i] <- datum[which(datum$days == 0 & datum$col_no == datum$col_no[i]), ]$infest_200
}

#    mites_tot
for(i in 1:nrow(datum)) {
  datum$pre_mites_tot[i] <- datum[which(datum$days == 0 & datum$col_no == datum$col_no[i]), ]$mites_tot
}

#    bees
for(i in 1:nrow(datum)) {
  datum$pre_bees[i] <- datum[which(datum$days == 0 & datum$col_no == datum$col_no[i]), ]$bees
}

#    bees_frames
for(i in 1:nrow(datum)) {
  datum$pre_bees_frames[i] <- datum[which(datum$days == 0 & datum$col_no == datum$col_no[i]), ]$bees_frames
}

#    wk_cells
for(i in 1:nrow(datum)) {
  datum$pre_wk_cells[i] <- datum[which(datum$days == 0 & datum$col_no == datum$col_no[i]), ]$wk_cells
}

#    wk_frames
for(i in 1:nrow(datum)) {
  datum$pre_wk_frames[i] <- datum[which(datum$days == 0 & datum$col_no == datum$col_no[i]), ]$wk_frames
}

#    honey_frames
for(i in 1:nrow(datum)) {
  datum$pre_honey_frames[i] <- datum[which(datum$days == 0 & datum$col_no == datum$col_no[i]), ]$honey_frames
}

#    pollen_frames
for(i in 1:nrow(datum)) {
  datum$pre_pollen_frames[i] <- datum[which(datum$days == 0 & datum$col_no == datum$col_no[i]), ]$pollen_frames
}

# Other variables that could be valuable predictors:
# dwv_num_sides_obs
# dwv_tf
# cdb_num_sides_obs


```

Save datum as a .csv
```{r}
write.csv(datum, "../2 Analysis/data_prepped/datum_prepped.csv")
```

